### 2015-10-17 Assignment I/Assignment1-Ditz_Schroeder-WS15-16.tex
alles mal auf den neuesten

%\textsl{}%!TEX TS-options = --shell-escape
%!TEX TS-program = pdflatex
\documentclass[%
   10pt,              % Schriftgroesse
   ngerman,           % wird an andere Pakete weitergereicht
   a4paper,           % Seitengroesse
   DIV11,             % Textbereichsgroesse (siehe Koma Skript Dokumentation !)
]{scrartcl}%     Klassen: scrartcl, scrreprt, scrbook, article
% -------------------------------------------------------------------------

\usepackage[utf8]{inputenc} % Font Encoding, benoetigt fuer Umlaute
\usepackage[english]{babel}   % \textsl{}Spracheinstellung

\usepackage[T1]{fontenc} % T1 Schrift Encoding
\usepackage{textcomp}    % Zusatzliche Symbole (Text Companion font extension)
\usepackage{lmodern,dsfont}     % Latin Modern Schrift
\usepackage{dsfont}
%\usepackage{wasysym}
\usepackage{ulem}
\usepackage{graphicx}
\usepackage{eurosym}
%\usepackage{txfonts}
\usepackage{stmaryrd}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{multirow}
\usepackage{listings}
\usepackage{etextools}
\usepackage{ifthen}
\usepackage{TikZ} %phylogenetischer Baum
\usetikzlibrary{calc, shapes, backgrounds} %für die Phylogenetische bäume
\usetikzlibrary{automata,arrows}
\usepackage{subfigure}

% Definition des Headers
\usepackage{geometry}
\geometry{a4paper, top=3cm, left=3cm, right=3cm, bottom=3cm, headsep=0mm, footskip=0mm}
\renewcommand{\baselinestretch}{1.3}\normalsize

\def\header#1#2#3#4#5#6#7{\pagestyle{empty}
\noindent
\begin{minipage}[t]{0.6\textwidth}
\begin{flushleft}
\textbf{#4}\\% Fach
#6\\% Semester
Tutor: #2  % Tutor 
\end{flushleft}
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\begin{flushright}
\points{#7}% Punktetabelle
\vspace*{0.2cm}
#5%  Names
\end{flushright}
\end{minipage}

\begin{center}
{\Large\textbf{ Blatt #1}} % Blatt

{(Abgabe am #3)} % Abgabedatum
\end{center}
}

\newenvironment{vartab}[1]
{
    \begin{tabular}{ |c@{} *{#1}{c|} } %\hline
}{
    \end{tabular}
}

\newcommand{\myformat}[1]{& #1}

\newcommand{\entry}[1]{
  \edef\result{\csvloop[\myformat]{#1}}
  \result \\ \hline
}

\newcommand{\numbers}[1]{
  \newcounter{ctra}
\setcounter{ctra}{1}
\whiledo {\value{ctra} < #1}%
{%
  \myformat{\thectra}
  \stepcounter{ctra}%
}
\myformat{\thectra}
}
\newcommand{\emptyLine}[1]{
  \newcounter{ctra1}
\setcounter{ctra}{1}
\whiledo {\value{ctra1} < #1}%
{%
  \myformat{\hspace*{0.5cm}}
  \stepcounter{ctra1}%
}
}

\newcommand{\points}[1]{
\newcounter{colmns}
\setcounter{colmns}{#1}
\stepcounter{colmns}
  \begin{vartab}{\thecolmns}
    \numbers{#1} & $\sum$\\\hline
    \emptyLine{\thecolmns}\\
  \end{vartab}
}

\begin{document}
%\header{Blatt}{Tutor}{Abgabedatum}{Vorlesung}{Bearbeiter}{Semester}{Anzahl Aufgaben}
\header{7}{Alexander Seitz}{19. October 2015}{Bioinformatics I}{\\Jonas Ditz \\\& Benjamin Schroeder}{WS 15/16}{3}

  \section*{Exercise 1 - \textsl{}}

  \section*{Exercise 2 - \textsl{DP Algorithm for finding motifs}}
  \emph{for} i = 0 \emph{to} \emph{length}(X)\\
  \indent \emph{if} X[i] == Y[0]\\
  \indent \indent Score = 0
  \indent \indent \emph{for} j = 0 \emph{to} \emph{length}(Y)\\
  \indent \indent \indent \emph{if} (X[i+j] == Y[j]) Score += 1\\
  \indent \indent if (Score > T) 
  \indent \indent \indent \emph{print}(motif found at position i)
  \indent \indent \indent i += \emph{length}(Y)

  \section*{Exercise 3 - \textsl{Programming assignment: Needleman-Wunsch alignment}}
  One task was to handle command line options. We wrote a new class to have a modular 
  and easy way to provide this functionality. The java package \textit{args4j} has everything
  we need. So we used that package to write our own option handler class. The code can 
  be found in \textit{CommandLineParser.java}. We did the same with gap penalties. A
  gap penalty class gives one the advantage of modular gap handling. If a gap penalties are
  needed, just reuse this class.\\
  A new member function of our NeedlemanWunsch class takes care of file writing. If a user 
  wants the resulting alignment in a FastA file, this function takes care of it.

\end{document}