### 2015-10-26 Assignment III/Assignment3-Ditz_Schroeder-WS15-16.tex
hier ist schonmal ein erster Entwurf für den Beweis in Aufgabe 1. Vielleicht kannst du mal drübergucken.

%\textsl{}%!TEX TS-options = --shell-escape
%!TEX TS-program = pdflatex
\documentclass[%
   10pt,              % Schriftgroesse
   nenglish,           % wird an andere Pakete weitergereicht
   a4paper,           % Seitengroesse
   DIV11,             % Textbereichsgroesse (siehe Koma Skript Dokumentation !)
]{scrartcl}%     Klassen: scrartcl, scrreprt, scrbook, article
% -------------------------------------------------------------------------

\usepackage[utf8]{inputenc} % Font Encoding, benoetigt fuer Umlaute
\usepackage[english]{babel}   % \textsl{}Spracheinstellung

\usepackage[T1]{fontenc} % T1 Schrift Encoding
\usepackage{textcomp}    % Zusatzliche Symbole (Text Companion font extension)
\usepackage{lmodern,dsfont}     % Latin Modern Schrift
\usepackage{dsfont}
%\usepackage{wasysym}
\usepackage{ulem}
\usepackage{graphicx}
\usepackage{grffile} %allows to use pngs
\usepackage{eurosym}
%\usepackage{txfonts}
\usepackage{stmaryrd}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{multirow}
\usepackage{listings}
\usepackage{etextools}
\usepackage{ifthen}
%\usepackage{TikZ} %phylogenetischer Baum
%\usetikzlibrary{calc, shapes, backgrounds} %für die Phylogenetische bäume
%\usetikzlibrary{automata,arrows}
\usepackage{subfigure}

% Definition des Headers
\usepackage{geometry}
\geometry{a4paper, top=3cm, left=3cm, right=3cm, bottom=3cm, headsep=0mm, footskip=0mm}
\renewcommand{\baselinestretch}{1.3}\normalsize

\def\header#1#2#3#4#5#6#7{\pagestyle{empty}
\noindent
\begin{minipage}[t]{0.6\textwidth}
\begin{flushleft}
\textbf{#4}\\% Fach
#6\\% Semester
Tutor: #2  % Tutor 
\end{flushleft}
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\begin{flushright}
\points{#7}% Punktetabelle
\vspace*{0.2cm}
#5%  Names
\end{flushright}
\end{minipage}

\begin{center}
{\Large\textbf{ Assignment #1}} % Blatt

{(Abgabe am #3)} % Abgabedatum
\end{center}
}

\newenvironment{vartab}[1]
{
    \begin{tabular}{ |c@{} *{#1}{c|} } %\hline
}{
    \end{tabular}
}

\newcommand{\myformat}[1]{& #1}

\newcommand{\entry}[1]{
  \edef\result{\csvloop[\myformat]{#1}}
  \result \\ \hline
}

\newcommand{\numbers}[1]{
  \newcounter{ctra}
\setcounter{ctra}{1}
\whiledo {\value{ctra} < #1}%
{%
  \myformat{\thectra}
  \stepcounter{ctra}%
}
\myformat{\thectra}
}
\newcommand{\emptyLine}[1]{
  \newcounter{ctra1}
\setcounter{ctra}{1}
\whiledo {\value{ctra1} < #1}%
{%
  \myformat{\hspace*{0.5cm}}
  \stepcounter{ctra1}%
}
}

\newcommand{\points}[1]{
\newcounter{colmns}
\setcounter{colmns}{#1}
\stepcounter{colmns}
  \begin{vartab}{\thecolmns}
    \numbers{#1} & $\sum$\\\hline
    \emptyLine{\thecolmns}\\
  \end{vartab}
}

\begin{document}
%\header{Blatt}{Tutor}{Abgabedatum}{Vorlesung}{Bearbeiter}{Semester}{Anzahl Aufgaben}
\header{3}{Alexander Seitz}{2. November 2015}{Bioinformatics I}{\\Jonas Ditz \\\& Benjamin Schroeder}{WS 15/16}{3}

\section*{Theoretical Assignment - \textsl{Equivalence of distance and similarity alignments}}
Assume there are two sequences of length n and m, respectively. The length of an alignment between
this two sequences is of length n+m. As it is written on the assignment following equation is 
valid:

\begin{equation}
 n + m = 2*M + \sum_k kg_k ,
\end{equation}

where M is the number of aligned characters. Since later one of our sequences is called a, we 
changed the latter a, which is used on the assignment sheet, to M. \\
Using this equation, we can write the distance of our two sequences (let us call them a and b) as

\begin{align}
 D(a,b) &= min\{\sum_M d(a,b) + \sum_k k g_k\} \nonumber\\
        &= min\{\sum_M c + \sum_k k g_k c/2 - \sum_M s(a,b) + \sum_k \hat{\gamma}(k) g_k\} \nonumber\\
        &= min\{c(n+m)/2 - \sum_M s(a,b) + \sum_k \hat{\gamma}(k) g_k\} \nonumber\\
        &= c(n+m)/2 - max\{\sum_M s(a,b) - \sum_k \hat{\gamma}(k) g_k\} \nonumber\\
        &= c(n+m)/2 - S(a,b)
\end{align}

Solving for S(a,b), we get:

\begin{equation}
 S(a,b) = c(n+m)/2 - D(a,b)
\end{equation}

So one can see that the Score is optimal if and only if the Distance is optimal.

\section*{Practical Assignment - \textsl{Using BLAT to align 454 reads to the Helicobacter pylori genome}}

\section*{Practical Assignment - \textsl{Bonus: Use SSAHA2 to align 535 reads to the Helicobacter pylori genome}}

\end{document}