### 2015-11-13 AssignmentV/Assignment5-Ditz_Schroeder-WS15-16.tex

%\textsl{}%!TEX TS-options = --shell-escape
%!TEX TS-program = pdflatex
\documentclass[%
   10pt,              % Schriftgroesse
   nenglish,           % wird an andere Pakete weitergereicht
   a4paper,           % Seitengroesse
   DIV11,             % Textbereichsgroesse (siehe Koma Skript Dokumentation !)
]{scrartcl}%     Klassen: scrartcl, scrreprt, scrbook, article
% -------------------------------------------------------------------------

\usepackage[utf8]{inputenc} % Font Encoding, benoetigt fuer Umlaute
\usepackage[english]{babel}   % \textsl{}Spracheinstellung

\usepackage[T1]{fontenc} % T1 Schrift Encoding
\usepackage{textcomp}    % Zusatzliche Symbole (Text Companion font extension)
\usepackage{lmodern,dsfont}     % Latin Modern Schrift
\usepackage{dsfont}
\usepackage{color}
%\usepackage{wasysym}
\usepackage{ulem}
\usepackage{graphicx}
\usepackage{grffile} %allows to use pngs
\usepackage{eurosym}
%\usepackage{txfonts}
\usepackage{stmaryrd}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{multirow}
\usepackage{listings}
\usepackage{etextools}
\usepackage{ifthen}
\usepackage{cite}
%\usepackage{TikZ} %phylogenetischer Baum
%\usetikzlibrary{calc, shapes, backgrounds} %für die Phylogenetische bäume
%\usetikzlibrary{automata,arrows}
\usepackage{subfigure}

% Definition des Headers
\usepackage{geometry}
\geometry{a4paper, top=3cm, left=3cm, right=3cm, bottom=3cm, headsep=0mm, footskip=0mm}
\renewcommand{\baselinestretch}{1.3}\normalsize

\def\header#1#2#3#4#5#6#7{\pagestyle{empty}
\noindent
\begin{minipage}[t]{0.6\textwidth}
\begin{flushleft}
\textbf{#4}\\% Fach
#6\\% Semester
Tutor: #2  % Tutor 
\end{flushleft}
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\begin{flushright}
\points{#7}% Punktetabelle
\vspace*{0.2cm}
#5%  Names
\end{flushright}
\end{minipage}

\begin{center}
{\Large\textbf{ Assignment #1}} % Blatt

{(Handed in #3)} % Abgabedatum
\end{center}
}

\newenvironment{vartab}[1]
{
    \begin{tabular}{ |c@{} *{#1}{c|} } %\hline
}{
    \end{tabular}
}

\newcommand{\myformat}[1]{& #1}

\newcommand{\entry}[1]{
  \edef\result{\csvloop[\myformat]{#1}}
  \result \\ \hline
}

\newcommand{\numbers}[1]{
  \newcounter{ctra}
\setcounter{ctra}{1}
\whiledo {\value{ctra} < #1}%
{%
  \myformat{\thectra}
  \stepcounter{ctra}%
}
\myformat{\thectra}
}
\newcommand{\emptyLine}[1]{
  \newcounter{ctra1}
\setcounter{ctra}{1}
\whiledo {\value{ctra1} < #1}%
{%
  \myformat{\hspace*{0.5cm}}
  \stepcounter{ctra1}%
}
}

\newcommand{\points}[1]{
\newcounter{colmns}
\setcounter{colmns}{#1}
\stepcounter{colmns}
  \begin{vartab}{\thecolmns}
    \numbers{#1} & $\sum$ (7)\\\hline   %add here Complete number of points ---------------------------
    \emptyLine{\thecolmns}\\
  \end{vartab}
}

\begin{document}
%\header{Blatt}{Tutor}{Abgabedatum}{Vorlesung}{Bearbeiter}{Semester}{Anzahl Aufgaben}
\header{5}{Alexander Seitz}{16. November 2015}{Bioinformatics I}{\\Jonas Ditz \\\& Benjamin Schroeder}{WS 15/16}{4}

\section*{Theoretical Assignment - \textsl{Comparison with at most $l$ mismatches}}

General we are now only looking at the worst case scenario, which means uniform distributed mistmatches in an alignment. Also we only look at $k>=1$, as $k=0$ is in our application not a useful result. Assume two sequences of length t with l mismatches.\\
 Than both sequences contain $l$ tuples of length $L = \lfloor\frac{t}{l+1}\rfloor$ and one tuple which has a length of maximal $\lfloor\frac{t}{l+1}\rfloor$ . \\
\\
\noindent So both sequences share $l$ or $l+1$ k-tuples, if $k = \lfloor\frac{t}{l+1}\rfloor$, and for each 
$k \le \lfloor\frac{t}{l+1}\rfloor$ they share $(l+1) * \lfloor\frac{L}{k}\rfloor$
k-tuples.

\section*{Theoretical Assignment - \textsl{Linear programming by hand}}

The feasible region of this linear program is shown in figure \ref{feasibleRegion}, where the red line 
is constraint 1, green line is constraint 2, constraint 3 is drawn as a blue line and the yellow line 
is constraint 4.

\begin{figure}[ht]
 \centering
 \includegraphics[width=0.7\textwidth]{img/figure_1.png}
 \caption{feasible region of linear program (grey area)}
 \label{feasibleRegion}
\end{figure}

\subsection*{solution if t=s=1}
Generally we only used integers in this task. The constraint for the algorithm are as following if t and s value is 1 now:

\begin{eqnarray}
max. 1*x_1+1*x_2 \label{con1}\\ 
s.t. 2*x_1+x_2 \le 12 \label{con2}\\ 
2*x_1+3*x_2 \le 7 \label{con3}\\ 
-1*x_1+2*x_2 \le 16 \label{con4}\\ 
x_1,x_2 \ge 0 \label{con5}
\end{eqnarray}

\noindent the next step we started to compute each line of the tuple. Starting with constraint \ref{con3} under the aspect of constraint \ref{con5}. All possible tuples respectively to both constraints from the assignment are:\\
$temp1=\{(0,7);(1,6);(2,5);(3,4);(4,3);(5,2);(6,1);(7,0)\}$\\

\noindent the constraint \ref{con2} was applied to temp1:\\
$temp2=\{(0,7);(1,6);(2,5);(3,4);(4,3);(5,2);\}$\\

\noindent the next constraint (constraint \ref{con4}) was applied:\\
$temp3=\{(5,2)\}$\\

\noindent but not least the last constraint (4) is checked and the remaining tuple from the last step is valid:\\
$r=\{(5,2)\}$\\

\noindent the final result for the Algorithm with s=t=1 is the tuple: (5,2). Solving the formula with this numbers 
leads us to the following maximal value: $1*x_1 + 1*x_2 = 5+2 = 7$\\

\subsection*{How to make unsolvable?}
To get a not solvable algorithm for example the following constraint could be added:
	\begin{equation}\\
		x_1+x_2\ge 7
	\end{equation}

\noindent a new constraint just needs to be a contradiction to another constraint.

\subsection*{How to get infinite solutions with t and s?}
The linear program has infinite solutions, if we assign the following values to s and t:

%Getting infinite solutions,  requires to describe with the max function of a linear algorithm a line in the feasible region and not just one point.
%Therefore we decided to use constraint 3 and wanted to generate a max function along constraint (3). For the modelling of the max function we used two points from constraint (3) ($P_1(2,4); P_2(5,2)$) to determine $s=\frac{2}{3}$ and $t=1$ in the following equations.

%\begin{eqnarray*}
%	s * x_{1_{P_1}} + t*x_{2_{P_1}} = s * x_{1_{P_2}} + t*x_{2_{P_2}}\\
%\Leftrightarrow	s*2+t*4 = s*5+ t*2\\
%\Leftrightarrow	2t = 3s\\
%\Leftrightarrow	s = \frac{2}{3}t\\
%\rightarrow t=1\\	
%\end{eqnarray*}

%Now we controlled the resuld by inserting the extreme data points $P_1$ and $P_2$.

%\begin{eqnarray*}
%t=1\\
%\frac{2}{3}*t*x_1+t*x_2\\
%\Leftrightarrow\frac{2}{3}*x_1+x_2\\
%\\
%\frac{2}{3}*2+4 =\frac{2}{3}*5+2  = 5+\frac{1}{3}\\
%\end{eqnarray*}

%Not serious ;)
%The equation influence by t and s is:

\begin{align}
 s &= x_2 \nonumber \\
 t &= -x_1 \nonumber \\
\end{align}

\noindent These values would lead to the equation $x_2*x_1 + (-x_1)*x_2$, which is always zero. So 
the maximal result is zero and you can fill in arbitrary numbers to get this result, i.e. infinite 
solutions.

\subsection*{how to make the points a optimal solution?}

For the first point $P_1=(0,3)$ we have to set $s = -1$ and $t = 1$. That result in a maximal value 
for (0,3) because we cannot set $x_2$ arbitrary big due to constraint 4. And if we assign a bigger 
value to $x_1$, would decrease the result. Also we would violate constraint 2, if $x_1$ is to big.\\
\\
P2=(2,4)\\

\begin{figure}[h]
	\includegraphics[width=\textwidth]{img/figure-P24.png}
	\label{P24}
\end{figure}

To achive that P(2,4) is the only maximal value, we need to use the borders used in figure \ref{P24}. The lower one is described by the maximal constraint (4) (equation \ref{lowerbound})and the higher one is described by the horizontal line at $x2=4$. Next we can define relations, for the variables $m_2$ and $c_2$ of the searched linear equation. These are shown in equation \ref{rel1} and \ref{rel2}.

\begin{eqnarray}
-1x_1+2x_2=6\\
\Leftrightarrow  x_2 = \frac{1}{2} x_1 +3 \label{lowerbound}\\
=	x_2= m_1 x_1 + c_1\\
	\\
	m_1 > m_2 > 0 \label{rel1}\\
	c_1<c_2<4 \label{rel2}\\
\end{eqnarray}

 So any linear equation, which lies in the green area of figure \ref{P24} , is a possible solution. We will now just choose one and define $c_2 = 3.5 = \frac{7}{2}$ according to relation \label{rel2}. From this now we can deduce $m_2=\frac{1}{4}$ as shown in Equation \ref{m} so that P(2,4) is the only intersection with the feasible region.

\begin{eqnarray}
4=m_2 * 2+3.5\\
m_2 = \frac{4-3.5}{2}  = \frac{1}{4} \label{m}\\
\end{eqnarray}

The last step is now to get s and t, this is achieved by forming the found linear equation back to the desired form (Equation \ref{result24}). The complete result is for $s=\frac{-1}{14}$ and for $t= \frac{2}{7}$.

\begin{eqnarray}
x_2=\frac{1}{4} * x_1 + \frac{7}{2}\\
 	\Leftrightarrow 	- \frac{1}{4} x_1 + x_2=\frac{7}{2}\\
 	\Leftrightarrow 	- \frac{1}{14} x_1 + \frac{2}{7} x_2 = 1 \label{result24}\\
\Rightarrow s= - \frac{1}{14} ; t=\frac{2}{7}\\
\end{eqnarray}

P3=(1,3)\\
\\
\noindent To make $P_4=(6,0)$ the maximum we set $s = 1$ and $t = -1$. We cannot set $x_1$ to a bigger value 
than 6 without violating constraint 1.

\section*{Theoretical Assignment - \textsl{Bonus: Carillo-Lipman bound}}

The Carrillo-Lipman bound is defined by $B_{pq} \coloneqq B(A) - (\sum_{i<j} s(\hat{A}_i, \hat{A}_j)-s(\hat{A}_p,\hat{A}_q))$, 
where $B(A)$ is a score for the alignment computed by a heuristic and $s(\hat{A}_i,\hat{A}_j)$ is 
the score of a optimal alignment of sequences $i$ and $j$. If we have just two sequences $p$ and $q$, 
this equation would be simplified as 
\[
B_{pq} = B(A) - (s(\hat{A}_p,\hat{A}_q)-s(\hat{A}_p,\hat{A}_q)) = B(A).
\]

\noindent So the Carrillo-Lipman bound for two sequences is just the computed score by a heuristic.
%\newpage
%\bibliography{Assignment4-Ditz_Schroeder-WS15-16.bib}
%\bibliographystyle{ieeetr}

\end{document}