### 2016-01-11 poodle_project/paper/main.tex

\documentclass{bioinfo}
\copyrightyear{2016} \pubyear{2016}

\access{}%Advance Access Publication Date: Day Month Year}
\appnotes{}%Manuscript Category}

\begin{document}
\firstpage{1}

\subtitle{}%Database applications}

\title[POODLE]{POODLE 1.0 - To the core of your data}
\author[Ditz, Schroeder]{Jonas Ditz\,$^{\text{\sfb 1,}*}$ and Benjamin Schroeder\,$^{\text{\sfb 1,}*}$} 
\address{$^{\text{\sf 1}}$Department of Informatics, Eberhard Karls University, T\"ubingen, 72076, Germany.}

\corresp{$^\ast$To whom correspondence should be addressed.}

\history{}%Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{}%Associate Editor: XXXXXXX}

\abstract{\textbf{Motivation:} Many life science laboratories are still using Excel files do organize their data. 
	This leads to a huge workload for maintenance as well as a inconvenient access and update routine. 
	POODLE provides an easy-to-use and powerful interface to improve the cloning work of your lab.\\
\textbf{Results:} POODLE is a Java-based web interface, which allows intuitive access, update and 
manipulation of data for cloning.\\
\textbf{Contact:} \href{}{forename.surname@student.uni-tuebingen.de}\\
\textbf{Supplementary information:} Supplementary data and the code are available at \textit{https://github.com/derjedi/BioinformaticsI\-Assigments/tree/master/poodle\_project}
online.}

\maketitle

\section{Introduction}

\enlargethispage{12pt}

\subsection{A real world pipeline}
function study
\subsection{cloning as lab strategy}
Toolkit idea
\subsection{cloning methods}
Classsic Cloning, Quickchange, Restriction free cloning

\subsection{The resulting data}
primer
cloningVectors
proteinConstructs

\section{Approach}

As already wrote above many life-science laboratories use Excel files for storing and working with 
data. But Excel files have a few drawbacks, which make working with them inconvenient. On of the 
biggest problems occurs, if more than one member of the laboratory tries to access the file at the 
same time. That can lead to inconsistent files. Second, searching in a Excel file is not as easy as 
it could be. Furthermore, licenses for Excel are expensive and free software is not as powerful as 
commercial ones. The reason for using Excel rather than a database system is that most life-scientist 
are not familiar with database systems and therefor chooses the GUI of Excel to work with. We are 
facing that problem by combining the functionality of a database system with an intuitive and 
easy-to-learn handling.

\begin{methods}
\section{Methods}

POODLE is build as a two-layer software. The first layer consists of the database and routines to 
automatically build and update that database. The second layer consists of the web service that is 
used to access, update and manipulate data. This is the front end layer and provides access to all 
functions for the user. Because POODLE is not just a data storage software but also comes with useful 
methods like a Blast search.

\subsection{Database layer}

SQLite\footnote{http://www.sqlite.org/} is the database system running in the background. We chose 
this software for several reasons. First, it is free of charge. Second, and more important SQLite is 
a small and fast database system written in C. So the requirement in space is very low. Since, the 
whole database is stored just in one file SQLite also has a incredibly good performance. And there 
is the possibility to create a mobile version of POODLE without changing the whole database system, 
which may be interesting in the future. Besides, 
SQLite guaranties that all transactions are ACID even if a system crash or power failure occurs. So 
we have a robust storage and access of data as well as a lot of functionality provided by the database 
system. In addition to the SQLite database POODLE's software creates a BLAST database for the provided 
data, too.

The build and update process is implemented in Python. So the user does not have to have knowledge in 
SQL coding but only has to execute one python script and everything is done automatically. This routine 
ensures that the SQLite database and the BLAST database are always in the same state. It also makes 
the database a little bit safer, since changes can be reviewed before inserting them into the database.

\subsection{Web service}

POODLE web service is build with Vaadin\footnote{https://vaadin.com/home}. We chose that framework 
because it allows a safe 
communication between the front end and the database layer as well as a smooth and easy handling of 
the side. Since a website developed with Vaadin is highly modular, it is easy to extend the web 
service in the future. That could be necessary, if new functionalities are required. Currently, 
there are three pages with different roles. The search page provides a formula that can be used 
to perform a simple SQL search in the local database. Therefor, different fields are displayed and 
can be filled by the user. A click on the search button will automatically build a SQL query and 
results are displayed, if there are available results. The new entry page allows to insert a new 
entry into the database. It also provide a Blast search against the NCBI databases to check, whether 
there are already available information about the new sequence or not. Last but not least, users can 
perform a local Blast search against their one database or a remote Blast search against the NCBI 
databases on the Blast page. That functionality can be found on the Blast page.

\subsection{Blast functionality}

Our software package provides not just a database functionality for powerful and secure data storage 
and a easy-to-use interface but also a Blast (\citealp{Altschul01}) functionality. This function 
is implemented by using a combination of Blast binaries provided by NCBI and the Biojava Api 
(\citealp{Prlic01}). There are two different Blast searches available in POODLE. On one hand, there 
is a remote Blast search on the NCBI server. And on the other hand, there is a local Blast search. 
The remote Blast search is not a fully functional copy of the NCBI web service but can used to 
perform a quick and simply search against databases provided by NCBI, e.g. Swissprot (\citealp{Doni01}). But we 
do not recommend to overuse that because too many requests may cause a blacklisting by NCBI. This 
remote Blast routine is mainly for the insertion of a new entry. The idea is to check whether there 
are information about the new sequence. But it can also used separated from the new entry routine. 
The local Blast routine does everything someone would 
expect from a Blast search but against the local database, hence local Blast. There a several programs, 
e.g. blastn or blastp, available for both Blast methods.

\end{methods}

\begin{figure}[!tpb]%figure1
\fboxsep=0pt\colorbox{gray}{\begin{minipage}[t]{235pt} \vbox to 100pt{\vfill\hbox to
235pt{\hfill\fontsize{24pt}{24pt}\selectfont FPO\hfill}\vfill}
\end{minipage}}
%\centerline{\includegraphics{fig01.eps}}
\caption{Caption, caption.}\label{fig:01}
\end{figure}

%\begin{figure}[!tpb]%figure2
%%\centerline{\includegraphics{fig02.eps}}
%\caption{Caption, caption.}\label{fig:02}
%\end{figure}

\section{Discussion}

POODLE is completely open-source. So the user can adapt the code at any point of the software. To make 
this adaption as easy as possible for the user we chose to use widely known software inside of POODLE. 
Most scientist have heard of SQL as a database managment system. So it is not to dangerous to assume 
that there are people with basic SQL knowledge in most of the laboratories. And that means these 
laboratories do not have to use POODLE as a black box. We chose SQLite over all other SQL databases 
for the reasons described in the method section.

The build up process of the database is written in Python. There are two reasons for that. First, 
Python is a easy and powerful language which makes it possible for the user to adapt the build up 
process without detailed programming knowledge. And second there is almost no 
Server without a Python2.7 or higher environment.

Vaadin was chose as the framework to be for several reasons. Java is a widely used programming language 
in the scientific world. So there is a relative high chance for a person in the laboratory, who speaks 
Java. Again we want to make adaption as easy as possible for the user. Furthermore, Vaadin needs a 
tomcat server. Such a server is provided free of charge by Apache\footnote{http://www.apache.org/}. 
So the user does not have to face extra charges just to be able to use POODLE. More reasons for 
Vaadin are discussed in the method section.

\section{Conclusion}

POODLE provides a secure storage of cloning data as well as a easy-to-use interface for 
interacting with these data. Besides, it comes with functionalities like local and remote Blast 
search that can improve the workflow during cloning.

It is our goal to improve the software in the future by adding new functionalities like a chromatogram 
viewer and the possibility to add annotations to sequences.

\section*{Acknowledgements}

We would like to thank Silke Wiesner and her group at the Max-Plack-Institute for Developmental Biology 
in T\"{u}bingen for providing us with real-world data and fruitful discussions.

\section*{Funding}

This work has been supported by ... nobody. Seriously, not anybody. We know that is sad but it is hard 
to convince people giving their money away.

%\bibliographystyle{natbib}
%\bibliographystyle{achemnat}
%\bibliographystyle{plainnat}
%\bibliographystyle{abbrv}
%\bibliographystyle{bioinformatics}
%
%\bibliographystyle{plain}
%
%\bibliography{Document}

\begin{thebibliography}{}

\bibitem[Altschul {\it et~al}., 1990]{Altschul01}
Altschul, S.F., Gish, W., Miller, W., Myers, E.W., Lipman, D.J. (1990) Basic local alignment search tool., {\it J. Mol. Biol.}, {\bf 215}, 403-410.

\bibitem[Prli\'{c} {\it et~al}., 2012]{Prlic01}
Prli\'{c}, A., Yates, A., Bliven, S. E., Rose, P. W., Jacobsen, J., Troshin, P. V., ... Willis, S. (2012) BioJava: an open-source framework for bioinformatics in 2012. {\it Bioinformatics}, {\bf 28}(20), 2693-2695. http://doi.org/10.1093/bioinformatics/bts494

\bibitem[O'Donavan {\it et~al}., 2002]{Doni01}
O'Donovan, C., Martin, M.J., Gattiker, A., Gasteiger, E., Bairoch, A., Apweiler, R. (2002) High-quality protein knowledge resource: SWISS-PROT and TrEMBL {\it Brief. Bioinform.}  {\bf 3} (3), 275-284. doi: 10.1093/bib/3.3.275

\end{thebibliography}
\end{document}