### 2016-01-20 AssignmentIX/Assignment11-Ditz_Schroeder-WS15-16.tex
erste aufgabe des neuen blatts ist fertig

%\textsl{}%!TEX TS-options = --shell-escape
%!TEX TS-program = pdflatex
\documentclass[%
   10pt,              % Schriftgroesse
   ngerman,           % wird an andere Pakete weitergereicht
   a4paper,           % Seitengroesse
   DIV11,             % Textbereichsgroesse (siehe Koma Skript Dokumentation !)
]{scrartcl}%     Klassen: scrartcl, scrreprt, scrbook, article
% -------------------------------------------------------------------------

\usepackage[utf8]{inputenc} % Font Encoding, benoetigt fuer Umlaute
\usepackage[english]{babel}   % \textsl{}Spracheinstellung

\usepackage[T1]{fontenc} % T1 Schrift Encoding
\usepackage{textcomp}    % Zusatzliche Symbole (Text Companion font extension)
\usepackage{lmodern,dsfont}     % Latin Modern Schrift
\usepackage{dsfont}
%\usepackage{wasysym}
\usepackage{ulem}
\usepackage{graphicx}
\usepackage{eurosym}
%\usepackage{txfonts}
\usepackage{stmaryrd}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{multirow}
\usepackage{listings}
\usepackage{etextools}
\usepackage{ifthen}
%\usepackage{TikZ} %phylogenetischer Baum
%\usetikzlibrary{calc, shapes, backgrounds} %für die Phylogenetische bäume
%\usetikzlibrary{automata,arrows}
\usepackage{subfigure}

% Definition des Headers
\usepackage{geometry}
\geometry{a4paper, top=3cm, left=3cm, right=3cm, bottom=3cm, headsep=0mm, footskip=0mm}
\renewcommand{\baselinestretch}{1.3}\normalsize

\def\header#1#2#3#4#5#6#7{\pagestyle{empty}
\noindent
\begin{minipage}[t]{0.6\textwidth}
\begin{flushleft}
\textbf{#4}\\% Fach
#6\\% Semester
Tutor: #2  % Tutor 
\end{flushleft}
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\begin{flushright}
\points{#7}% Punktetabelle
\vspace*{0.2cm}
#5%  Names
\end{flushright}
\end{minipage}

\begin{center}
{\Large\textbf{ Blatt #1}} % Blatt

{(Abgabe am #3)} % Abgabedatum
\end{center}
}

\newenvironment{vartab}[1]
{
    \begin{tabular}{ |c@{} *{#1}{c|} } %\hline
}{
    \end{tabular}
}

\newcommand{\myformat}[1]{& #1}

\newcommand{\entry}[1]{
  \edef\result{\csvloop[\myformat]{#1}}
  \result \\ \hline
}

\newcommand{\numbers}[1]{
  \newcounter{ctra}
\setcounter{ctra}{1}
\whiledo {\value{ctra} < #1}%
{%
  \myformat{\thectra}
  \stepcounter{ctra}%
}
\myformat{\thectra}
}
\newcommand{\emptyLine}[1]{
  \newcounter{ctra1}
\setcounter{ctra}{1}
\whiledo {\value{ctra1} < #1}%
{%
  \myformat{\hspace*{0.5cm}}
  \stepcounter{ctra1}%
}
}

\newcommand{\points}[1]{
\newcounter{colmns}
\setcounter{colmns}{#1}
\stepcounter{colmns}
  \begin{vartab}{\thecolmns}
    \numbers{#1} & $\sum$ (10)\\\hline
    \emptyLine{\thecolmns}\\
  \end{vartab}
}

\begin{document}
%\header{Blatt}{Tutor}{Abgabedatum}{Vorlesung}{Bearbeiter}{Semester}{Anzahl Aufgaben}
\header{11}{Alexander Seitz}{18. January 2016}{Bioinformatics I}{\\Jonas Ditz \\\& Benjamin Schroeder}{WS 15/16}{2}

\section*{Theoretical Assignment - \textit{Assembly using de Bruijn graph}}

The de Bruijn graph for the reads provided on the assignment sheet would look like the graph in
Figure \ref{fig:4mere}, if one chooses $k = 4$. As one can see easily, there are not just one but 
four different Eulerian paths in the graph. Each path results in one of the following superstring:

\begin{align}
 S_1 &= AACCGTTAACGTAAACGT \nonumber \\
 S_2 &= AACCGTAAACGTTAACGT \nonumber \\
 S_3 &= AACCGTTAAACGTAACGT \nonumber \\
 S_4 &= AACCGTAACGTTAAACGT \nonumber
\end{align}

This happens due to the fact that there are nodes with more than one outgoing an incoming edge. If
one chooses $k = 5$ the resulting de Bruijn graph would look like the graph in Figure \ref{fig:5mere}.
It is obvious that the number of nodes is depended on $k$. With a bigger $k$ there are more nodes in 
the resulting de Bruijn graph. But in this case there is one problem. Since there are more than two 
unbalanced nodes in the de bruijn graph there is no Eulerian path.

\begin{figure}[h]
 \centering
 \includegraphics[width=0.7\textwidth]{deBruijnGraph_4mere.png}
 \caption{De Bruijn graph for the given reads and $k = 4$.}
 \label{fig:4mere}
\end{figure}

\begin{figure}[h]
 \centering
 \includegraphics[width=1\textwidth]{deBruijnGraph_5mere.png}
 \caption{De Bruijn graph for the given reads and $k = 5$.}
 \label{fig:5mere}
\end{figure}

\section*{Practical Assignment - \textit{Assemble the human mitochondrial genome}}

\end{document}